Project Manager's Progress Report

### 2014-05-17 pnsac/newsletter/article_EditorsNotes.tex
Updated files for Apr/May 2014 newsletter.

This issue of the Chronicle carries the last of four chapters in
Richard Lodge's story of his time at Rolls-Royce. Taken together, they
make an interesting ethnology of an extinct industrial culture. The
whole series is accessible through our website,
{\color{blue}\texttt{\url{http://www.projectnorthstar.ca}}} (note that
the location of the newsletter archive will change when the new
website goes live) starting with the April 2013 issue of the
Chronicle.

@@

We also have another hair-raising flying story from Tim Timmins. A
series of previous Timmins stories can be found through the same
website, starting with the December 2010 issue.

@@

Motorheads among our readers should look into Ted Devey's extensive
writings on the Rolls-Royce Merlin, starting in the December 2006
issue.

@@

All together, the archived issues are a valuable web resource for
readers with an interest in aviation history or the history of Project
North Star.

>>>>

While our project has been on hold, your editor still had to find some
content for the Chronicle.

%This issue of the Chronicle carries the last of four chapters in
%Richard Lodge's story of his time at Rolls-Royce. Taken together, they
%make an interesting ethnology of an extinct industrial culture. The
%whole series is accessible through our website,
%{\color{blue}\texttt{\url{http://www.projectnorthstar.ca}}} (note that
%the location of the newsletter archive will change when the new
%website goes live) starting with the April 2013 issue of the
%Chronicle.

@@

Jim Riddoch came to the rescue with a story about a
successful\/disastrous flight test.  We like these kinds of
stories--to small and personal to turn up in the history books, but
they do pull you into real human experiences. There will be more of these.

%We also have another hair-raising flying story from Tim Timmins. A
%series of previous Timmins stories can be found through the same
%website, starting with the December 2010 issue.

@@

Your editor became a writer too, and the writer submitted a tragic
story about a French aviatice. The story was accepted by the editor in
a blatant conflict of interest.

%Motorheads among our readers should look into Ted Devey's extensive
%writings on the Rolls-Royce Merlin, starting in the December 2006
%issue.

@@

And we bring you a poem by David Lambeth, an ode to the North Star
that was first published in the Chronicle of November 2005. Volunteer
Tim Timmins had found it in a hidey-hole in the aeroplane in the early
days of our restoration project.  Enjoy!

%All together, the archived issues are a valuable web resource for
%readers with an interest in aviation history or the history of Project
%North Star.

### 2014-05-17 pnsac/newsletter/article_Feature2.tex
Updated files for Apr/May 2014 newsletter.

\title{Building troop seats for North Star 17515}

\author{Rolf Geiger}

@@

There are no original parts or drawings for the seats. Fortunately,
the tubing and brackets that connect the seats to the aircraft wall
were still in place in the aircraft. Also, two lower tubing support
brackets were found hiding behind the galley, allowing us to produce
exact copies of the original. Everything else concerning the seats is
missing and needs to  be replaced.

\end{multicols}

%\begin{figure*}[htbp]
\begin{figure*}[h]
   \vspace{2em}
   \centering
   %name of the graphic, without the path AND in EPS format:
   \includegraphics[scale=0.75]{ass.eps}
   %caption of the figure 
   \caption*{\small \em Troop seat construction detail.}
   %label of the figure, which has to correspond to \ref{}:
   \label{fig:ass}
\end{figure*}

\begin{multicols}{2}

To find the size and the number of seats to be produced we measured
the space between the anchoring studs in the aircraft floor, which the
legs snap onto. This gave us the width of the seats. Thus, we will
produce (5) three-seaters plus (6) four-seaters, providing seating for
39 troops. No litters are planned.

Mike Irvin (our Museum coordinator) decided to use seats from a
helicopter in the Museum collection as an example for our seats. We
are more or less copying the parts of these seats, considering that we
have to adapt our methods to the equipment available to us.

To start the project, drawings of the parts were made. The list of
materials of these parts calls mainly for aluminum bar stock, aluminum
tubing, steel tubing, aluminum sheet and steel sheet.

\begin{figure}[htbp]
   \vspace{2em}
   \centering
   %name of the graphic, without the path AND in EPS format:
   \includegraphics[scale=0.5]{rolf.eps}
   %caption of the figure 
   \caption*{\small \em Rolf Geiger at work on the troop seats.}
   %label of the figure, which has to correspond to \ref{}:
   \label{fig:rolf}
\end{figure}

Thus, we need to produce (15) different metal parts, including: lower
tubing plus support brackets and clamps, spreader bars with two shell
clamps each to clamp onto the lower support tubing plus the front
brackets holding the spreader bars to the frontal support tubes, leg
brackets connecting the legs to the frontal support tubes plus
secondary brackets (slipping over the leg brackets) holding a spring
loaded pin with which the seat fabric will be tautened, legs,
including the lower clamping mechanism fastening the legs to the
floor. Most parts need (22) of each, except the shell clamps for the
spreader bars (44), hooks (117), leg grippers (44).

The procedure in the production of these parts is as with any project:
Take the finished product and then work backwards to determine the
production stages.

In the case of a formed sheet metal part a drawing of a flattened
blank is produced by calculating bend radii and adding that to the
straight surfaces (in effect un-bending or stretching the finished
part).

Experimental operations may be needed to find the shape and dimensions
of the blanks.  Tooling may be needed to produce the formed sheet
metal parts.  The blanks are made by shearing and sawing, according
to the parts drawing.

In the case of a machined part, it is usually produced from bar stock:
Parts are cut to length and subsequently machined as per
drawing. Partial in-between drawings may be necessary to produce the
parts in progressive stages.

To avoid the higher costs associated with surface plating the metal
parts, (as per original parts) we chose painting them instead, which
will be acceptable given the limited use of the finished product.

>>>>

\title{The Whispering Giant}
\author{Dave Lambeth}

@@

Twinkle, twinkle, old North Star,\\
With wrinkled skin and creaky
spar,\\
For years you’ve always been on
sked,\\
As many of our force you’ve sped\\
to distant lands beyond the skies,\\
you’ve never failed to see the rise\\
of distant suns or moonrise glare.\\
You’ve carried us with tender care\\
-back you came to take us home,\\
from old Japan or France or Rome\\
from England, Goose, and Kef,\\
We cannot hear you.\\
(‘cause we’re deaf)\\

### 2014-05-17 pnsac/newsletter/article_ProgReport.tex
Updated files for Apr/May 2014 newsletter.

\title{Project Manager's Progress Report}
\subtitle{December 2013}
\author{Bruce Gemmill}

@@

@@

Engine 3 is nearing completion.  The supercharger and intercooler
preheat assemblies have been installed, along with the propeller
reduction gearbox.  While this work was being carried out by the
engine crew, the remainder of the volunteer work force busied
themselves completing the numerous cowl panels, pipes and hoses and
other pieces needed to complete the engine.  Two items that seemed to
take forever were the large steel exhaust shrouds.  A lot of hammering
was needed to pound out years of dents, then numerous rivets and all
the Dzus fasteners and springs needed to be replaced.  The complex
shape of the exhaust shroud made this challenging work.  Further
delaying completion was the search for suitable clear coat to protect
the steel from corrosion.  The old clearcoat we used successfully on
engines 1 and 2 is no longer available.  Mike Irvin finally located a
suitable replacement, and the shrouds were clear coated and serial
numbers were stencilled on.

@@

The auxiliary gearbox was disassembled, and significant corrosion was
found on the bearings inside.  Bearings had to be ordered from
England.  Meanwhile, the gearbox was cleaned and painted.  At the same
time, the large DC generator was dismantled, repainted and
reassembled, as well as the air pump and the tachometer generator and
propeller synchronizer.  Once the new bearings were installed, the
gearbox was completed and the generator and air pump were attached.
Finally, the gearbox was installed on the firewall, which is now
finished, awaiting the installation of engine \#3.

@@

Work has slowed on the forward section of the aircraft.  The repair
needed under the floor of the crew lounge is on hold until a suitable
piece of extruded aluminum can be found.  Until this is complete, the
new cushion flooring cannot be installed. This has also delayed the
installation of all of the crew lounge accessories, such as the table,
seats and secure storage bin, all of which are complete but waiting in
storage.

\section{Fuselage and Empennage}
\label{empenage}

The forward and rear belly compartments are now complete.  All the
frame pieces and hardware needed for the troop seats have been made,
and work is progressing on the seats and backs.  One section has
undergone a trial fit, and all pieces seem to fit perfectly.

\section{Planned Restoration Work 2013/14}
\label{sec:plannedwork}

The museum has advised all North Star volunteers that our work must be
put on hold for several months while a major exhibit for the 100th
anniversary of the beginning of the First World War is prepared.  Our
work space in the restoration hangar is needed to assemble and prepare
several aircraft that will be added to the display space in March.
Until then our work on the aircraft is suspended, but we don't expect
this delay to have a serious impact on the overall project.  We look
forward to resuming work in the Spring.

>>>>

\title{Project Manager's Progress Report}
\subtitle{April 2014}
\author{Bruce Gemmill}

@@

Due to museum preparations for a display to mark the beginning of
the First World War and other events in 2014, the restoration
space used by Project North Star was taken over by other museum
projects, and most work on the North Star was suspended.
However, some limited work did continue.

@@

Work on Engine 3 was halted for several weeks in early 2014, but
has now resumed on a reduced work schedule.  Most accessories are
now installed, and soon the cowl panels will be fitted.  It is
expected that Engine 3 will be installed late this spring or
early summer, so that work can begin on removing and restoring
Engine 4.

This downtime also offered the opportunity for the engine crew to
clean and re-organize the engine shop.

@@

\section{Cockpit}
\label{cockpit}

While restoration work in the shop could not be continued, we
used the opportunity to install the crew seats and other restored
items to the cockpit.  The pilot seat was installed (first
officer seat was installed last year) then the railings and
safety board were installed behind the seats.  Next we installed
the radio officer and navigator tables, and the seats for these
positions.  The radio officer seat proved to be a particular
challenge, as the floor in this area is no longer flat, and some
modifications to the seat rails were necessary to allow the seat
to slide.  Finally the flight engineer’s jump seat was attached
to the railing.  With five crew seats installed, the cockpit is
very cramped.

The cockpit installations were finished off by installing the
curtains, crew oxygen hoses, and a Morse code key at the radio
position.

@@

The floor under this section was repaired by removing a section
of badly corroded “I” beam and fabricating a new one from several
pieces of aluminum which were riveted in place to support the
floor.  Once this repair was completed, new cushion flooring was
cut and glued to the crew lounge and the washroom floor, then
various pieces of floor trim were installed.  To install the crew
seat, it was necessary to remove the lower support section, since
the fully assembled seat could not easily fit through the narrow
doorway.  Once the seat was installed, we added the crew table
and bunk bed, along with the secure storage box.  We also
installed the curtain rod and new curtains to close off the bunk
bed. The installation was finished off by re-installing the door
between the lounge and the main cabin, so that this fully
restored area can be protected from dust and debris while the
main cabin undergoes restoration over the next two years.

The completion of the forward area of the aircraft marks a
significant milestone in the restoration of the North Star.

\section{Planned Restoration Work 2014}
\label{sec:plannedwork}

Some limited work is being done to remove fittings and wall
panels from the main cabin, in preparation for some major
restoration work this summer.  Fabric work on the troop seats is
nearing completion, although the seats will not be installed
until all work on the main cabin is complete.  We hope to resume
a full work schedule by June, with the aircraft outside once
more.
